%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% acknowledgements.tex
%% NOVA thesis document file
%%
%% Text with acknowledgements
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgements

The work presented in this document would never see the light of day if not for the collaboration of several people to whom I wish to manifest my profound gratitude and recognition. 

First I would like to thank my advisors and members of the iCBD Project, Professors Paulo Lopes, Pedro Medeiros and Nuno Preguiça, for there advice, always supporting this work and the countless hours spent trying to find the better course of action especially when I would see myself lost and overwhelmed.

It is necessary to give a special thanks to Engineer Miguel Martins from Reditus S.A, to whom I had the pleasure of work closely for more than a year. Who taught me a lot, not only from his vast knowledge of computer systems, backed by countless years of experience in the IT world and without who this work would not be possible but also for the many conversations we held in the realms of Physics, Economics and History. I will always admire how a person can keep such a large body of knowledge and a passion for sharing it.

I also like to extend my recognition to Dr. Henrique Mamede and Engineer Sérgio Rita, also from Reditus S.A., for the opportunity given, constant support and the warm-hearted welcome in my year-long stay with SolidNetworks. And, to Luís Anjos from the \textit{Divisão de Infraestruturas Informáticas da FCT NOVA} (Div-I FCT NOVA) for all his help in resolving issues related to FCT NOVA's internal network.

Finally, my very heartfelt gratitude to my family and friends for their unconditional support, for putting up with my grumbles when work problems went home with me, being always there with a kind word and some crazy plan to make me forget work for a couple of hours and enjoy their friendship and time spent together.

I also would like to acknowledge the following institutions for their hosting and financial support: \textit{Departamento de Informática and Faculdade de Ciências e Tecnologia of the Universidade NOVA de Lisboa} (DI-FCT NOVA); the \textit{NOVA Laboratory of Computer Science and Informatics} (NOVA LINCS) in particular the Computer Systems group; \textit{SolidNetworks – Business Consulting, Lda} of the \textit{Reditus S.A. Group}; and the funding provided through the \textit{COMPETE2020 / PORTUGAL2020} program for the \textit{iCBD} project (POCI-01-0247-FEDER-011467).